\Large Plotting the Single-Electron Solution to Schr\"{o}dinger's Equation with OpenCL

### 2010-05-08 paper/paper.tex
Paper

@@

@@

Once computation has completed, the buffer is copied back from working memory (either video memory or system memory) to the final image buffer in system memory. Once there, pixel values are scaled linearly so that the brightest pixel is 100\% white.

@@

@@

\begin{figure}
    \includegraphics[width=84.5mm]{overheadPlot.pdf}
    \caption{Overhead from image buffer copy as a percentage of the total runtime, lower numbers are better}
    \label{fig:overheadPlot}
\end{figure}

@@

We've also found that --- for problems which are embarrassingly easy to parallelize, like this one --- it's very easy to implement your problem using OpenCL. We've used MPI and pthreads in the past, and it seems that OpenCL (perhaps because it's a more modern API) is a bit easier to use in terms of implementation. The added benefit of kernels running on a wider array of hardware (not just the CPU, like MPI, but on the GPU as well) means that OpenCL is a no-brainer --- at least for this sort of simple project. It's quite likely that a problem requiring a large amount of inter-kernel communication would be better suited by something with more communications primitives, like MPI.

>>>>

Quantum mechanics --- since Louis de Broglie's paper {\it Research on Quantum Theory} --- tells us that matter behaves sometimes as a wave, and other times as a particle. The electron in a hydrogen atom is no different: when it isn't being observed, it doesn't exist at a particular point in space, but instead expands as a wave. The atomic orbital cloud which we're plotting here is actually simply a plot of the density of that wave at all points surrounding the atom's nucleus.

@@

Indeed, while all of the benchmarks in this paper were done with $\psi_c$ included, one might notice that --- since it's a constant term across the entire image, and we're scaling the image by {\it another} constant scale factor --- that it doesn't actually affect the visualization, and can be entirely discarded. It appears that discarding the overhead required to pass this value in as an argument to each of the kernel instances and to multiply it into the value of $\psi$ gives a performance increase of between 6 and 10 percent: quite a significant improvement for no change in output!

@@

Once computation has completed, the buffer is copied back from working memory (either video memory or system memory) to the final image buffer in system memory. Once there, pixel values are scaled linearly so that the brightest pixel is 100\% white, and the image is compressed to PNG and saved to the filesystem.

@@

\begin{figure}
    \includegraphics[width=84.5mm]{overheadPlot.pdf}
    \caption{Overhead from image buffer copy as a percentage of the total runtime; smaller values are better}
    \label{fig:overheadPlot}
\end{figure}

@@

@@

We've also found that --- for problems which are embarrassingly easy to parallelize, like this one --- it's very easy to implement your problem using OpenCL. We've used MPI and pthreads in the past, and it seems that OpenCL (perhaps because it's a more modern API) is a bit easier to use in terms of implementation. The added benefit of kernels running on a wider array of hardware (not just the CPU, like MPI, but on the GPU as well) means that OpenCL is a no-brainer --- at least for this sort of simple project. It's quite likely that a problem requiring a large amount of inter-kernel communication would be better suited by something with more communications primitives, like MPI. Indeed, besides {\it barrier}, OpenCL doesn't seem to have any manner of inter-kernel communications functions.
